---
title: "Analysis of The Characteristics of the Oldest Living Cats"
always_allow_html: true
author: "Rachel Bellflowers"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    keep_md: true
bibliography: references.bib  
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 6, fig.height = 6)

#knitr::write_bib(c("knitr", "plotly", "xml2", "rvest", "tidyverse", "ggthemes", "DT", "data.table", "kableExtra"), "references.bib")

library(xml2)
library(rvest)
library(tidyverse)
library(plotly)
library(ggthemes)
library(data.table)
library(DT)
library(knitr)
library(kableExtra)

cats <- "https://en.wikipedia.org/wiki/List_of_oldest_cats"

table <- read_html(cats) %>%
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[5]/div[1]/table[2]') %>%
  html_table(fill = TRUE)

table <- table[[1]]

table <- table[-29, -c(9, 11:12)]

table[table == ""] <- NA

colnames(table)[5] <- "reported_age"

table[5] <- c("38.0082191780822", "38", "36.0027397260274", "34.4109589041096", "34.1643835616438", "33.4986301369863", "33", "33", "31.1643835616438", "32", "31.9150684931507", "31.1643835616438", "31", "30.1123287671233", "30", "30", "29", "29", "28", "28", "28", "27.1917808219178", "27", "27", "27", "27", "26.0712328767123", "26")

table$reported_age <- as.numeric(table$reported_age)

```

---

# Introduction

This is an analysis of an unofficial list from [Wikipedia](https://en.wikipedia.org/wiki/List_of_oldest_cats) of the world's oldest cats. In order for cats to be included onto this list, they must have lived to the minimum age of 25 years old. Despite this dataset being small and imprecise, I hope in analyzing it to discover possible factors contributing to these cats' longevity. 

---

# Method

To obtain this dataset, I used the `rvest` package to webscrape the table from its Wikipedia page. In the case of cats with inexact ages, I calculated their approximate age. For instance, if a cat was listed as being 29 years old and 8 months, I inputted their age as $29 + (8 * 30 )/365 = 29.6575$ days. I then used the `plotly` package to create visualizations to examine the dataset. Finally, I conducted several t tests to determine any significant differences between variables that might be linked to age at time of death. 

---

# Results

---

## Exploratory Analysis

```{r} 
mean <- mean(table$reported_age)
sd <- sd(table$reported_age)

mean <- round(mean, 4)
sd <- round(sd, 4)
```


The ages of the cats ranged from 25 to 38, with a mean of `r mean` and a standard deviation of `r sd`. The majority of the cats resided in the United Kingdom. There were 15 males, 12 females, and one cat of unknown sex that was removed from the dataset. The breeds of the cats included domestic shorthairs, a Maine Coon, a Siamese, a Sphynx-Devon Rex, and others identified by the color of their coat.

<br>

```{r ages}
tab_age <- table %>%
  mutate(Name = fct_reorder(Name, reported_age)) 

ages <- plot_ly(tab_age, x = ~Name, y = ~reported_age, type = "bar")

ages <- ages %>% layout(title = "Reported Ages", yaxis = list(title = "Reported age"))
                        
ages
```


<br><br>


```{r country}

country <- plot_ly(table, x = ~Country, type = "histogram")
 
country <- country %>% layout(title = "Country of Origin")
 
country
```

<br>
<br>

```{r breeds}
breed <- plot_ly(table, x = ~Breed, type = "histogram")

breed <- breed %>% layout(title = "Breeds")

breed
```

<br>
<br>


```{r sex}

sex <- plot_ly(table, x = ~Sex, type = "histogram")

sex <- sex %>% layout(title = "Count of Sex")

sex
```


---

```{r include = FALSE}

# Making sex numeric

table$sex_num <- table$Sex
table$sex_num[which(table$Sex == "female")] <- "1"
table$sex_num[which(table$Sex == "male")] <- "0"
table$sex_num <- as.numeric(table$sex_num)

# Making country numeric

table$country_num <- table$Country
table$country_num[which(table$Country == "Australia")] <- "0"
table$country_num[which(table$Country == "Canada")] <- "1"
table$country_num[which(table$Country == "New Zealand")] <- "2"
table$country_num[which(table$Country == "Sweden")] <- "3"
table$country_num[which(table$Country == "Switzerland")] <- "4"
table$country_num[which(table$Country == "United Kingdom")] <- "5"
table$country_num[which(table$Country == "United States")] <- "6"
table$country_num <- as.numeric(table$country_num)

# Preparing columns for t tests


males <- table$reported_age[which(table$sex_num == 0)]
females <- table$reported_age[which(table$sex_num == 1)]



aus <- table$reported_age[which(table$Country == "Australia")]
us <- table$reported_age[which(table$Country == "United States")]
uk <- table$reported_age[which(table$Country == "United Kingdom")]
nz <- table$reported_age[which(table$Country == "New Zealand")]
swit <- table$reported_age[which(table$Country == "Switzerland")]
swed <- table$reported_age[which(table$Country == "Sweden")]
can <- table$reported_age[which(table$Country == "Canada")]


# T tests

var.test(males, females) # > 0.05
t.test(males, females, var.equal = TRUE) # t = 0.51665, p = 0.6099

var.test(us, uk) # > 0.05
t.test(us, uk, var.equal = TRUE) # t = 0.34046, p = 0.7371

var.test(us, can) # < 0.05
t.test(us, can, var.equal = FALSE) # t = 2.0606, p = 0.07329

var.test(uk, can) # < 0.05
t.test(uk, can, var.equal = FALSE) # t = 3.1346, p = 0.008618
```

## Testing Differences Between Groups

I ran t tests on the variables sex and country of origin. Only the countries of Canada, the United Kingdom, and the United States of America had a sufficient number of values for these tests. The only significant result ($p < 0.05$) found was between the UK and Canada, with cats in the former country living slightly longer (30.6032 years on average) compared to those living in Canada (28 years on average).

```{r results='asis'}
variables <- c("sex", "usa_uk", "usa_canada", "uk_canada")
t_scores <- c(0.5167, 0.3405, 2.0606, 3.1346)
p_values <- c(0.6099, 0.7371, 0.0733, 0.0086)

t_tests <- data.table(variables, t_scores, p_values)

values_table <- kable(t_tests, caption = "T Test Scores")

values_table %>%
  row_spec(4, background = "#66ff66") %>%
  kable_material()
```
---

## Discussion

This study suffered from several major issues. First, with a sample size of 27, it is likely that the results found here do not adequately represent the target population. Secondly, given that many of the subjects were described by their appearnace rather than by their breed, I decided against running any analytics on that variable. In the future, I would recommend attempting to determine other factors that were not universally provided in the Wikipedia article, such as level of exercise, diet, and breed, that would be helpful in understanding this dataset better.


---

# References
