---
title: "Analysis of The Characteristics of the Oldest Living Cats"
always_allow_html: true
author: "beep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    keep_md: true
bibliography: references.bib  
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#knitr::write_bib(c("knitr", "plotly", "xml2", "rvest", "tidyverse", "ggthemes", "DT", "data.table", "kableExtra"), "references.bib")

library(xml2)
library(rvest)
library(tidyverse)
library(plotly)
library(ggthemes)
library(data.table)
library(DT)
library(knitr)
library(kableExtra)

cats <- "https://en.wikipedia.org/wiki/List_of_oldest_cats"

table <- read_html(cats) %>%
  html_nodes(xpath = '/html/body/div[3]/div[3]/div[5]/div[1]/table[2]') %>%
  html_table(fill = TRUE)

table <- table[[1]]

table <- table[-29, -c(9, 11:12)]

table[table == ""] <- NA

colnames(table)[5] <- "reported_age"

table[5] <- c("38.0082191780822", "38", "36.0027397260274", "34.4109589041096", "34.1643835616438", "33.4986301369863", "33", "33", "31.1643835616438", "32", "31.9150684931507", "31.1643835616438", "31", "30.1123287671233", "30", "30", "29", "29", "28", "28", "28", "27.1917808219178", "27", "27", "27", "27", "26.0712328767123", "26")

table$reported_age <- as.numeric(table$reported_age)

```

---

# Introduction

This is an analysis of an unofficial list from [Wikipedia](https://en.wikipedia.org/wiki/List_of_oldest_cats) of world's oldest cats. The reported ages range from precise to vague estimates. The minimum age requirement for cats to be added to this list was 25 years old. Despite this dataset being small and imprecise, I hope by examining it to discover possible factors contributing to these cats' longevity. 

---

# Method

To obtain this dataset, I used the `rvest` package to webscrape the table from its Wikipedia page. In the case of cats with imprecise ages, I calculated their approximate age. For instance, if a cat was listed as being 29 years old and 8 months, I inputted their age as $29 + (8 * 30 )/365 = 29.6575$ days. I then used the `plotly` package to create visualizations to examine the dataset. Finally, I conducted several t tests to determine any significant differences between variables that might be linked to age at time of death. 

---

# Results

---

## Exploratory Analysis

```{r} 
mean <- mean(table$reported_age)
sd <- sd(table$reported_age)

mean <- round(mean, 4)
sd <- round(sd, 4)
```


The ages of the cats ranged from 25 to 38, with a mean of `r mean` and a standard deviation of `r sd` . As can be seen in Chart 2, the majority of cats resided in the United Kingdom. There were 15 males, 12 females, and one cat of unknown sex that was removed from the dataset.

<br>

```{r}
tab_age <- table %>%
  mutate(Name = fct_reorder(Name, reported_age)) 

ages <- plot_ly(tab_age, x = ~Name, y = ~reported_age, type = "bar")

ages <- ages %>% layout(title = "Ages and Names", yaxis = list(title = "Reported age"))
                        
ages
```


<br><br>


```{r country}

country <- plot_ly(table, x = ~Country, type = "histogram")
 
country <- country %>% layout(title = "Country of Origin")
 
country
```

<br>
<br>

```{r breed}
breed <- plot_ly(table, x = ~Breed, type = "histogram")

breed <- breed %>% layout(title = "Cat Breeds")

breed
```

<br>
<br>


```{r sex}

sex <- plot_ly(table, x = ~Sex, type = "histogram")

sex <- sex %>% layout(title = "Count of Sex")

sex
```


---

```{r include = FALSE}

# Making sex numeric

table$sex_num <- table$Sex
table$sex_num[which(table$Sex == "female")] <- "1"
table$sex_num[which(table$Sex == "male")] <- "0"
table$sex_num <- as.numeric(table$sex_num)

# Making country numeric

table$country_num <- table$Country
table$country_num[which(table$Country == "Australia")] <- "0"
table$country_num[which(table$Country == "Canada")] <- "1"
table$country_num[which(table$Country == "New Zealand")] <- "2"
table$country_num[which(table$Country == "Sweden")] <- "3"
table$country_num[which(table$Country == "Switzerland")] <- "4"
table$country_num[which(table$Country == "United Kingdom")] <- "5"
table$country_num[which(table$Country == "United States")] <- "6"
table$country_num <- as.numeric(table$country_num)

# Preparing columns for t tests


males <- table$reported_age[which(table$sex_num == 0)]
females <- table$reported_age[which(table$sex_num == 1)]



aus <- table$reported_age[which(table$Country == "Australia")]
us <- table$reported_age[which(table$Country == "United States")]
uk <- table$reported_age[which(table$Country == "United Kingdom")]
nz <- table$reported_age[which(table$Country == "New Zealand")]
swit <- table$reported_age[which(table$Country == "Switzerland")]
swed <- table$reported_age[which(table$Country == "Sweden")]
can <- table$reported_age[which(table$Country == "Canada")]


# T tests

var.test(males, females) # > 0.05
t.test(males, females, var.equal = TRUE) # t = 0.51665, p = 0.6099

var.test(us, uk) # > 0.05
t.test(us, uk, var.equal = TRUE) # t = 0.34046, p = 0.7371

var.test(us, can) # < 0.05
t.test(us, can, var.equal = FALSE) # t = 2.0606, p = 0.07329

var.test(uk, can) # < 0.05
t.test(uk, can, var.equal = FALSE) # t = 3.1346, p = 0.008618
```

## Testing Differences Between Groups


```{r results='asis'}
variables <- c("sex", "usa_uk", "usa_canada", "uk_canada")
t_scores <- c(0.5167, 0.3405, 2.0606, 3.1346)
p_values <- c(0.6099, 0.7371, 0.0733, 0.0086)

t_tests <- data.table(variables, t_scores, p_values)

values_table <- kable(t_tests, caption = "T Test Scores")

values_table %>%
  row_spec(4, background = "#66ff66") %>%
  kable_material()
```
---

## Discussion

---

# References
